CF Registry Analysis

### 2025-08-05 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
CFRD diagnosis comparisons for A1c registry paper

```{r}
# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
# This is super slow but idc idc
first_cfrd=head(first_cfrd,100)
t <- apply(first_cfrd, 1, function(r) {
  id <- as.numeric(r["eDWID"])
  year <- as.numeric(r["first_cfrd"])
  d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
  if (any(!is.na(d$A_hgba1c))) {
    a1c <- T
  } else {
    a1c <- F
  }
  if (any(!is.na(d["A_twohour_bloodglucose"]) |
    !is.na(d["A_ogttfast_bloodglucose"]))) {
    ogtt <- T
  } else {
    ogtt <- F
  }
  if (a1c & ogtt) {
    return("both")
  } else if (a1c & !ogtt) {
    return("a1c")
  } else if (ogtt & !a1c) {
    return("ogtt")
  } else if (!ogtt & !a1c) {
    return("neither")
  }
})
first_cfrd$a1c_ogtt <- t
cfrd_only <- left_join(first_cfrd, annual |> remove_labels(),
  by = join_by(eDWID, first_cfrd == ReviewYear)
)
cfrd_only$a1c_ogtt <- factor(cfrd_only$a1c_ogtt,
  levels = c("both", "ogtt", "a1c", "neither"),
  labels = c("OGTT", "OGTT", "HbA1c Only", "Neither")
)
cfrd_only %>%
  select(
    a1c_ogtt, Age_YrEnd, A_weight, A_height, A_bmivalue, A_bmipercentile,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, Race, Hispanicrace, MutClass,
    A_hgba1c, A_ogttfast_bloodglucose, A_twohour_bloodglucose, A_IsOnEnzymes,
    liver_disease, A_cultureresults3
  ) %>%
  tbl_summary(by = a1c_ogtt, missing_text = "Missing") %>%
  add_overall() %>%
  add_p(test=list(all_continuous() ~ "oneway.test"))
```

@@

```{r results='asis'}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
  group_by(eDWID) %>%
  summarise(
    `Age Group` = last(na.omit(age_group)),
    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose))
  )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(ogtts$OGTTs,
  levels = 0:3, labels = c(
    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
  )
)
# Make table for each variable
age <- ogtts %>%
  select(`Age Group`, OGTTs) %>%
  tbl_summary(by = `Age Group`) %>%
  add_ci() %>%
  add_p() %>%
  print()
gender <- ogtts %>%
  select(Gender, OGTTs) %>%
  tbl_summary(by = Gender) %>%
  add_ci() %>%
  add_p() %>%
  print()
muts <- ogtts %>%
  select(MutClass, OGTTs) %>%
  tbl_summary(by = MutClass) %>%
  add_ci() %>%
  add_p() %>%
  print()
race <- ogtts %>%
  select(Race, OGTTs) %>%
  tbl_summary(by = Race) %>%
  add_ci() %>%
  add_p() %>%
  print()
eth <- ogtts %>%
  select(Hispanicrace, OGTTs) %>%
  tbl_summary(by = Hispanicrace) %>%
  add_ci() %>%
  add_p() %>%
  print()
```

>>>>

# Diagnosis comparisons for HbA1c registry paper

```{r}
# Per reviewer, out of X individuals with a diagnosis of CFRD, X number had an
# OGTT (and X had an A1c, but no OGTT) that year or the year prior.
# This is super slow but idc idc
t <- apply(first_cfrd, 1, function(r) {
  id <- as.numeric(r["eDWID"])
  year <- as.numeric(r["first_cfrd"])
  d <- annual[annual$eDWID == id & annual$ReviewYear %in% c(year, year - 1), ]
  if (any(!is.na(d$A_hgba1c))) {
    a1c <- T
  } else {
    a1c <- F
  }
  if (any(!is.na(d["A_twohour_bloodglucose"]) |
    !is.na(d["A_ogttfast_bloodglucose"]))) {
    ogtt <- T
  } else {
    ogtt <- F
  }
  if (a1c & ogtt) {
    return("both")
  } else if (a1c & !ogtt) {
    return("a1c")
  } else if (ogtt & !a1c) {
    return("ogtt")
  } else if (!ogtt & !a1c) {
    return("neither")
  }
})
first_cfrd$a1c_ogtt <- t
cfrd_only <- left_join(first_cfrd, annual %>% remove_labels(),
  by = join_by(eDWID, first_cfrd == ReviewYear)
)
cfrd_only$a1c_ogtt <- factor(cfrd_only$a1c_ogtt,
  levels = c("both", "ogtt", "a1c", "neither"),
  labels = c("OGTT", "OGTT", "HbA1c Only", "Neither")
)
cfrd_only %>%
  select(
    a1c_ogtt, Age_YrEnd, A_weight, A_height, A_bmivalue, A_bmipercentile,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, Race, Hispanicrace, MutClass,
    A_hgba1c, A_ogttfast_bloodglucose, A_twohour_bloodglucose, A_IsOnEnzymes,
    liver_disease, A_cultureresults3
  ) %>%
  tbl_summary(
    by = a1c_ogtt, missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) %>%
  add_overall() %>%
  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
  modify_caption("**All Groups**")
cfrd_only %>%
  filter(a1c_ogtt %in% c("OGTT", "HbA1c Only")) %>%
  mutate(a1c_ogtt = droplevels(a1c_ogtt)) %>%
  select(
    a1c_ogtt, Age_YrEnd, A_weight, A_height, A_bmivalue, A_bmipercentile,
    A_FEV1_pct_predicted, A_FVC_pct_predicted, Race, Hispanicrace, MutClass,
    A_hgba1c, A_ogttfast_bloodglucose, A_twohour_bloodglucose, A_IsOnEnzymes,
    liver_disease, A_cultureresults3
  ) %>%
  tbl_summary(
    by = a1c_ogtt, missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) %>%
  add_overall() %>%
  add_p(test = list(all_continuous() ~ "oneway.test")) %>%
  modify_caption("**OGTT vs. HbA1c Only**")
```

@@

```{r results='asis'}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
  group_by(eDWID) %>%
  summarise(
    `Age Group` = last(na.omit(age_group)),
    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) |
      !is.na(A_ogttfast_bloodglucose))
  )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(ogtts$OGTTs,
  levels = 0:3, labels = c(
    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
  )
)
# Make table for each variable
age <- ogtts %>%
  select(`Age Group`, OGTTs) %>%
  tbl_summary(by = `Age Group`) %>%
  # add_ci() %>%
  add_p() %>%
  print()
gender <- ogtts %>%
  select(Gender, OGTTs) %>%
  tbl_summary(by = Gender) %>%
  # add_ci() %>%
  add_p() %>%
  print()
muts <- ogtts %>%
  select(MutClass, OGTTs) %>%
  tbl_summary(by = MutClass) %>%
  # add_ci() %>%
  add_p() %>%
  print()
race <- ogtts %>%
  select(Race, OGTTs) %>%
  tbl_summary(by = Race) %>%
  # add_ci() %>%
  add_p() %>%
  print()
eth <- ogtts %>%
  select(Hispanicrace, OGTTs) %>%
  tbl_summary(by = Hispanicrace) %>%
  # add_ci() %>%
  add_p() %>%
  print()
```
